---
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(datamuseum)
```

`datamuseum` is an R package designed to improve specimen management and
improve access to legacy databases. This extends from first steps, such as
removing samples with unclear or missing labels, to aesthetic improvements
intended for publishable tables and figures.

`datamuseum` is categorized into four main parts, and each will be described
briefly here. Please refer to our
[Articles](https://btorgovitsky00.github.io/datamuseum/articles/)
for more in-depth workflows and breakdowns.

# Geographic Data

`datamuseum` includes a host of functions designed for manipulating and
refining geographic data. Although the package as a whole is intended for
biological samples, functions in the `latlong_*` group are universally
applicable.

Management with `datamuseum` is done in part with
[`rnaturalearth`](https://CRAN.R-project.org/package=rnaturalearth),
which provides a standard reference for geography and works alongside other
functions of the package to ensure mappable data is consistent and ready for
display with other packages like
[`ggplot2`](https://CRAN.R-project.org/package=ggplot2).

# Taxonomic Data

At its core, `datamuseum` is intended for use with biological specimen
data sets, with a focus on taxonomy. `datamuseum` accomplishes the task
of robust taxonomic review in its taxon function group through a 
dual-reference system which checks against the Global Biodiversity 
Information Facility's rGBIF package, and the Integrated Taxonomic Information 
System via taxize. Users have the option to use GBIF, ITIS, or both for their
taxonomic inquiries.

`datamuseum` is only as good as its sources, and user input is occasionally
essential. Although its ability to double-check provides safety against misspellings 
or inconsistent updates, some cases may still arise where manual changes are
still necessary.

# Example Data

In lieu of its limitations, `datamuseum` includes practice data in the form of
accessioned octopus specimen data sets refined to Japan. Superfamily Octopodoidea was 
selected and refined to Japan due to the long history of octopus fisheries and cuisine 
in the region, as well as recent regional taxonomic updates. 

More specifically, *Octopus vulgaris* (Cuvier, 1797) in East Asia was recently
re-described as the also valid *Octopus sinensis* (d'Orbigny, 1834) by
[Gleadall et al. (2016)](https://www.jstage.jst.go.jp/article/specdiv/21/1/21_31/_article).
`datamuseum` would struggle to reflect a regional change like this since both
species are still valid and accepted by GBIF and ITIS. As a result, we
recommend adding a simple command like the one below to your `datamuseum`
workflow for case-specific changes. 

This code, and its broader workflow with
associated `taxon_*` and `latlong_*` functions, can also be found at
[Octopodoidea in Japan](https://btorgovitsky00.github.io/datamuseum/articles/octopodoidea_japan.html).

```{r Sinensis -> Vulgaris, eval = FALSE}
museum_clean <- museum_clean %>% 
  mutate(SciName = case_when(
    SciName == "Octopus vulgaris"        ~ "Octopus sinensis",
    TRUE                                 ~ SciName
  ))
```

# Utilities

Going beyond manipulating just geographic and taxonomic data, `datamuseum`
also provides management options for accessioned data sets. This includes
options for removing repeat data (deduplicating) when multiple sources
are accessioned, expanding your data set when one specimen lot contains
multiple individuals, and preparing your taxonomic data for presentation on
a ggplot2 graph!

```{r logo, include = FALSE, echo = FALSE, eval = FALSE}
library(hexSticker)
library(ggplot2)
library(png)
library(grid)

img <- readPNG("D:/TU_Research/TU Dissertation Chapter 1/dataMuseum/man/figures/sophie_logo.png")

p <- ggplot() +
  annotation_raster(img, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  coord_fixed(ratio = nrow(img) / ncol(img)) +
  theme_void() +
  theme(panel.background = element_rect(fill = "transparent", color = NA),
        plot.background  = element_rect(fill = "transparent", color = NA))

sticker(
  subplot  = p,
  package  = "",
  p_size   = 0,
  p_color  = "#FFFFFF",
  h_fill   = "#c2edff",
  h_color  = "#003580",
  s_x      = 1,
  s_y      = 1,
  s_width  = 1.75,
  s_height = 1.45,
  white_around_sticker = FALSE,
  filename = "D:/TU_Research/TU Dissertation Chapter 1/dataMuseum/man/figures/logo.png"
)
```

```

